---
title: "Rainfall"
author: "CHEN XIN WEI 17202767"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

## Get data

```{r setup, include=FALSE}
getwd()
Rainfall <- read.csv("Rainfall.csv")
```

## Data Exploration

```{r}
names(Rainfall)
class(Rainfall)
str(Rainfall)
dim(Rainfall)
summary(Rainfall)
head(Rainfall)
tail(Rainfall)
```

**First Variable Drop**

```{r}
Rainfall <- subset(Rainfall, select = -c(name,datetime,precipprob,preciptype,snow,snowdepth,sunrise,sunset,conditions,stations))
str(Rainfall)
```

**Transform categorical variable**

```{r}
Rainfall$description<-as.factor(Rainfall$description)
Rainfall$icon<-as.factor(Rainfall$icon)
Rainfall$uvindex<-as.factor(Rainfall$uvindex)
```

**Checking relationship between cat var with dependent var**

```{r}
corr<-aov(precip ~ description * uvindex * icon, data=Rainfall)
summary(corr)
```

**correlation matric**

```{r}
Rainfall_sub <- subset(Rainfall, select = -c(description,icon,uvindex))
library(corrplot)
cr_Rainfall <- cor(Rainfall_sub)
cr_Rainfall
corrplot(cr_Rainfall,method="color")
```

## Second Variable Drop

```{r}
Rainfall <- subset(Rainfall, select = -c(windgust, windspeed, winddir, sealevelpressure, moonphase, severerisk, icon))
str(Rainfall)
```

## Check missing value

```{r}
sum(is.na(Rainfall))
colSums(is.na(Rainfall))
```

## Check outliers

```{r}
boxplot(Rainfall$tempmax)
boxplot(Rainfall$tempmin)
boxplot(Rainfall$temp)
boxplot(Rainfall$feelslikemax)
boxplot(Rainfall$feelslikemin)
boxplot(Rainfall$feelslike)
boxplot(Rainfall$dew)
boxplot(Rainfall$humidity)
boxplot(Rainfall$precip)
boxplot(Rainfall$precipcover)
boxplot(Rainfall$cloudcover)
boxplot(Rainfall$visibility)
boxplot(Rainfall$solarradiation)
boxplot(Rainfall$solarenergy)
```

**Check duplication**

```{r}
sum(duplicated(Rainfall))
```

**skewness**

```{r}
hist(Rainfall$tempmax)
hist(Rainfall$tempmin)
hist(Rainfall$temp)
hist(Rainfall$feelslikemax)
hist(Rainfall$feelslikemin)
hist(Rainfall$feelslike)
hist(Rainfall$dew)
hist(Rainfall$humidity)
hist(Rainfall$precip)
hist(Rainfall$precipcover)
hist(Rainfall$cloudcover)
hist(Rainfall$visibility)
hist(Rainfall$solarradiation)
hist(Rainfall$solarenergy)
```

**skewness test**

```{r}
install.packages("moments")
library(moments)
skewness(Rainfall$tempmax,na.rm=TRUE)
skewness(Rainfall$tempmin,na.rm=TRUE)
skewness(Rainfall$temp,na.rm=TRUE)
skewness(Rainfall$feelslikemax,na.rm=TRUE)
skewness(Rainfall$feelslikemin,na.rm=TRUE)
skewness(Rainfall$feelslike,na.rm=TRUE)
skewness(Rainfall$dew,na.rm=TRUE)
skewness(Rainfall$humidity,na.rm=TRUE)
skewness(Rainfall$precip,na.rm=TRUE)
skewness(Rainfall$precipcover,na.rm=TRUE)
skewness(Rainfall$cloudcover,na.rm=TRUE)
skewness(Rainfall$visibility,na.rm=TRUE)
skewness(Rainfall$solarradiation,na.rm=TRUE)
skewness(Rainfall$solarenergy,na.rm=TRUE)
```

**Categorical variable**

```{r}
library(ggplot2)
ggplot(Rainfall,aes(x=uvindex))+ geom_bar()
ggplot(Rainfall,aes(x=description))+ geom_bar()
```

**correlation**

```{R}
Rainfall_sub <- subset(Rainfall, select = -c(description, uvindex))
cr_Rainfall <- cor(Rainfall_sub)
cr_Rainfall
corrplot(cr_Rainfall,method="color")
```

## Bivariate Analysis

```{r}
library(tidycomm)
correlate(Rainfall)
```

```{r}
correlate(Rainfall)%>% 
  to_correlation_matrix()
```

```{r}
plot(Rainfall$tempmax, Rainfall$precip, main="Scatter Plot with Max Temperature", xlab="Max Temperature", ylab="Precipitation", col="blue", pch=16)
abline(lm(Rainfall$precip ~ Rainfall$tempmax), col="red")
```

```{r}
plot(Rainfall$tempmin, Rainfall$precip, main="Scatter Plot with Min Temperature", xlab="Min Temperature", ylab="Precipitation", col="blue", pch=16)
abline(lm(Rainfall$precip ~ Rainfall$tempmin), col="red")
```

```{r}
plot(Rainfall$temp, Rainfall$precip, main="Scatter Plot with Temperature", xlab="Temperature", ylab="Precipitation", col="blue", pch=16)
abline(lm(Rainfall$precip ~ Rainfall$temp), col="red")
```

```{r}
plot(Rainfall$feelslikemax, Rainfall$precip, main="Scatter Plot with Max Temperature Feel", xlab="Max Temperature Feel", ylab="Precipitation", col="blue", pch=16)
abline(lm(Rainfall$precip ~ Rainfall$feelslikemax), col="red")
```

```{r}
plot(Rainfall$feelslikemin, Rainfall$precip, main="Scatter Plot with Min Temperature Feel", xlab="Min Temperature Feel", ylab="Precipitation", col="blue", pch=16)
abline(lm(Rainfall$precip ~ Rainfall$feelslikemin), col="red")
```

```{r}
plot(Rainfall$feelslike, Rainfall$precip, main="Scatter Plot with Temperature Feel", xlab="Temperature Feel", ylab="Precipitation", col="blue", pch=16)
abline(lm(Rainfall$precip ~ Rainfall$feelslike), col="red")
```

```{r}
plot(Rainfall$dew, Rainfall$precip, main="Scatter Plot with Dew", xlab="Dew", ylab="Precipitation", col="blue", pch=16)
abline(lm(Rainfall$precip ~ Rainfall$dew), col="red")
```

```{r}
plot(Rainfall$humidity, Rainfall$precip, main="Scatter Plot with Humidity", xlab="Humidity", ylab="Precipitation", col="blue", pch=16)
abline(lm(Rainfall$precip ~ Rainfall$humidity), col="red")
```

```{r}
plot(Rainfall$precipcover, Rainfall$precip, main="Scatter Plot with Precipitation Cover", xlab="Precipitation Cover", ylab="Precipitation", col="blue", pch=16)
abline(lm(Rainfall$precip ~ Rainfall$precipcover), col="red")
```

```{r}
plot(Rainfall$cloudcover, Rainfall$precip, main="Scatter Plot with Cloud Cover", xlab="Cloud Cover", ylab="Precipitation", col="blue", pch=16)
abline(lm(Rainfall$precip ~ Rainfall$cloudcover), col="red")
```

```{r}
plot(Rainfall$visibility, Rainfall$precip, main="Scatter Plot with Visibility", xlab="Visibility", ylab="Precipitation", col="blue", pch=16)
abline(lm(Rainfall$precip ~ Rainfall$visibility), col="red")
```

```{r}
plot(Rainfall$solarradiation, Rainfall$precip, main="Scatter Plot with Solar Radiation", xlab="Solar Radiation", ylab="Precipitation", col="blue", pch=16)
abline(lm(Rainfall$precip ~ Rainfall$solarradiation), col="red")
```

```{r}
plot(Rainfall$solarenergy, Rainfall$precip, main="Scatter Plot with Solar Energy", xlab="Solar Energy", ylab="Precipitation", col="blue", pch=16)
abline(lm(Rainfall$precip ~ Rainfall$solarenergy), col="red")
```

```{r}
corr<-aov(precip ~ description * uvindex, data=Rainfall)
summary(corr)
```

## **Multivariate Analysis**

**Multiple Linear Regression**

```{r}
Precipitation <- lm(precip ~ description + uvindex + tempmax + tempmin + temp + feelslikemax + feelslikemin + feelslike + dew + humidity + precipcover + cloudcover + visibility + solarradiation + solarenergy, data = Rainfall)
anova(Precipitation)
```

```{r}
summary(Precipitation)
```

```{r}
# Get the model residuals
model_residuals = Precipitation$residuals

# Plot the result
hist(model_residuals)
```

```{r}
# Plot the residuals
qqnorm(model_residuals)
# Plot the Q-Q line
qqline(model_residuals)
```

```{r}
#Install and Load necessary libraries 
install.packages("caret") 
install.packages("nnet") 
install.packages("randomForest") 
install.packages("Metrics") 
install.packages("ggplot2")

library(caret) 
library(nnet) 
library(randomForest) 
library(Metrics) 
library(ggplot2)
```

```{r}
#Load dataset and standardize time format for easy sorting
data<- read.csv("Rainfall1.csv") 
data$datetime <- as.Date(data$datetime, format="%m/%d/%Y") 
head(data) 
data <- data[order(data$datetime), ] 
data
```

# Split the dataset into features and target variable

```{r}
features <- data[, c("tempmax", "tempmin", "temp", "dew", "humidity", "precipcover", "windspeed", "winddir", "cloudcover", "visibility")] 
target <- data$precip
```

# Calculate the index for an 80-20 split

```{r}
set.seed(123) split_index <- floor(0.8 * nrow(data)) split_index
```

# Split the data into training and testing sets based on the index

```{r}
train_features <- features[1:split_index, ] 
train_target <- target[1:split_index] 
test_features <- features[(split_index + 1):nrow(data), ] 
test_target <- target[(split_index + 1):nrow(data)]
```

# Standardize the features

```{r}
preproc <- preProcess(train_features, method = c("center", "scale")) train_features_scaled <- predict(preproc, train_features) 
test_features_scaled <- predict(preproc, test_features)
```

# Combine scaled features with target for training

```{r}
train_data <- data.frame(train_target, train_features_scaled) 
names(train_data)[1] <- "precip"
```

# Train the ANN model

```{r}
ann_model <- nnet(precip ~ ., data = train_data, size = 10, linout = TRUE, maxit = 150)
```

# Predict on test set

```{r}
ann_predictions <- predict(ann_model, test_features_scaled) 
# Ensure that all predictions are at least 0 
ann_predictions <- ifelse(ann_predictions < 0, 0, ann_predictions)
```

# Calculate accuracy metrics

```{r}
ann_mae <- mae(test_target, ann_predictions) 
ann_mse <- mse(test_target, ann_predictions) 
ann_rmse <- sqrt(ann_mse) 
ann_r2<-R2(test_target, ann_predictions) 
print(paste("ANN Model MAE:", ann_mae)) 
print(paste("ANN Model MSE:", ann_mse)) 
print(paste("ANN Model RMSE:", ann_rmse)) 
print(paste("ANN Model R2:", ann_r2))
```

# Plot the actual vs predicted values for the test set

```{r}
test_datetime <- data$datetime[801:1000] ggplot() + geom_line(aes(x = test_datetime, y = test_target), color = "blue") + geom_line(aes(x = test_datetime, y = ann_predictions), color = "red") + ggtitle("ANN Model - Actual vs Predicted Precipitation") + xlab("Date") + ylab("Precipitation") 

#END OF ANN
```

```{r}
#RF 
data1 <- read.csv("Rainfall1.csv") data1$datetime <- as.Date(data1$datetime, format="%m/%d/%Y") data1 <- data1[order(data1$datetime), ]
```

# Split the dataset into features and target variable

```{r}
features1 <- data1[, c("tempmax", "tempmin", "temp", "dew", "humidity", "precipcover", "windspeed", "winddir", "cloudcover", "visibility")] 
target1 <- data1$precip 
train_features1 <- features1[1:split_index, ] 
train_target1 <- target1[1:split_index] 
test_features1 <- features1[(split_index + 1):nrow(data1), ] 
test_target1 <- target1[(split_index + 1):nrow(data1)]
```

# Train the RF model

```{r}
train_data1 <- data.frame(train_target1, train_features1) 
names(train_data1)[1] <- "precip" 
rf_model <- randomForest(precip ~ ., data = train_data1, ntree = 500)
```

# Predict on test set

```{r}
rf_predictions <- predict(rf_model, test_features1)
```

# Ensure that all predictions are at least 0

```{r}
rf_predictions <- ifelse(rf_predictions < 0, 0, rf_predictions)
```

# Calculate accuracy metrics

```{r}
rf_mae <- mae(test_target1, rf_predictions) 
rf_mse <- mse(test_target1, rf_predictions) 
rf_rmse <- sqrt(rf_mse) 
rf_r2<-R2(test_target1, rf_predictions) 
print(paste("RF Model MAE:", rf_mae)) 
print(paste("RF Model MSE:", rf_mse)) 
print(paste("RF Model RMSE:", rf_rmse)) 
print(paste("RF Model R2:", rf_r2)) 
# Plot the actual vs predicted values for the test set 
test_datetime_rf <- data1$datetime[(split_index + 1):nrow(data1)] ggplot() + geom_line(aes(x = test_datetime_rf, y = test_target), color = "blue") + geom_line(aes(x = test_datetime_rf, y = rf_predictions), color = "green") + ggtitle("RF Model - Actual vs Predicted Precipitation") + xlab("Date") + ylab("Precipitation")

#Multilinear Regression
data2 <- read.csv("Rainfall1.csv") 
data2$datetime <- as.Date(data1$datetime, format="%m/%d/%Y") 
data2 <- data2[order(data1$datetime), ]
```

# Split the dataset into features and target variable

```{r}
features2 <- data2[, c("tempmax", "tempmin", "temp", "dew", "humidity", "precipcover", "windspeed", "winddir", "cloudcover", "visibility")] 
target2 <- data2$precip 
train_features2 <- features2[1:split_index, ] 
train_target2 <- target2[1:split_index] 
test_features2 <- features2[(split_index + 1):nrow(data1), ] 
test_target2 <- target2[(split_index + 1):nrow(data1)] 
# Train the LM model 
train_data2 <- data.frame(train_target2, train_features2) 
names(train_data2)[1] <- "precip" 
lm_model <- lm(precip ~ ., data = train_data2)
```

# Predict on test set

```{r}
lm_predictions <- predict(lm_model, newdata = test_features2)
```

# Ensure that all predictions are at least 0

```{r}
lm_predictions <- ifelse(lm_predictions < 0, 0, lm_predictions)
```

# Calculate accuracy metrics

```{r}
lm_mae <- mae(test_target2, lm_predictions) 
lm_mse <- mse(test_target2, lm_predictions) 
lm_rmse <- sqrt(lm_mse) 
lm_r2<-R2(test_target2, lm_predictions) 
print(paste("Linear Regression Model MAE:", lm_mae)) 
print(paste("Linear Regression Model MSE:", lm_mse)) 
print(paste("Linear Regression Model RMSE:", lm_rmse)) 
print(paste("Linear Regression Model R2:", lm_r2))
```

# Plot the actual vs predicted values for the test set

```{r}
ggplot() + geom_line(aes(x = test_datetime, y = test_target2), color = "blue") + geom_line(aes(x = test_datetime, y = lm_predictions), color = "red") + ggtitle("Linear Regression Model - Actual vs Predicted Precipitation") + xlab("Date") + ylab("Precipitation")
```

```{r}
#SVM 
library(e1071)

#insert data and order
data4 <- read.csv("Rainfall1.csv") 
data4$datetime <- as.Date(data4$datetime, format="%m/%d/%Y") 
data4 <- data4[order(data4$datetime), ]
```

# Split the dataset into features and target variable

```{r}
features4 <- data4[, c("tempmax", "tempmin", "temp", "dew", "humidity", "precipcover", "windspeed", "winddir", "cloudcover", "visibility")] 
target4 <- data4$precip 
train_features4 <- features4[1:split_index, ] 
train_target4 <- target4[1:split_index] 
test_features4 <- features4[(split_index + 1):nrow(data4), ] 
test_target4 <- target4[(split_index + 1):nrow(data4)]

#SVM modelling 
train_data4 <- data.frame(train_target4, train_features4) 
names(train_data4)[1] <- "precip" 
svm_model <- svm(precip ~ ., data = train_data4, kernel = "linear", cost = 10) svm_predictions <- predict(svm_model, newdata=test_features4)
```

# Ensure that all predictions are at least 0

```{r}
svm_predictions <- ifelse(svm_predictions < 0, 0, svm_predictions)
```

# Calculate accuracy metrics

```{r}
svm_mae <- mae(test_target4, svm_predictions) 
svm_mse <- mse(test_target4, svm_predictions) 
svm_rmse <- sqrt(svm_mse) 
svm_r2<-R2(test_target4, svm_predictions) 
print(paste("svm Model MAE:", svm_mae)) 
print(paste("svm Model MSE:", svm_mse)) 
print(paste("svm Model RMSE:", svm_rmse)) 
print(paste("svm Model R2:", svm_r2))

#plot 
test_datetime_svm <- data4$datetime[(split_index + 1):nrow(data4)] ggplot() + geom_line(aes(x = test_datetime_svm, y = test_target4), color = "blue") + geom_line(aes(x = test_datetime_svm, y = svm_predictions), color = "green") + ggtitle("Support Vector Machine - Actual vs Predicted Precipitation") + xlab("Date") + ylab("Precipitation")

#Evaluation Table
Model_evaluation <- data.frame( model_name=c("Artificial Neutral Network Model","Random Forest","Multiple Linear Regression","Support Vector Machine"), mae_value=c("10.345","7.839","9.722","4.547"), rmse_value=c("18.669","11.823","12.938","9.204"), R2_score=c("0.114","0.157","0.169","0.157") ) 
print(Model_evaluation)
```

# Combine predictions into a DataFrame

```{r}
combined_predictions <- data.frame( ANN_Predictions = ann_predictions, RF_Predictions = rf_predictions, LM_Predictions = lm_predictions, SVM_Predictions = svm_predictions )
```

# Write the DataFrame to a CSV file

```{r}
write.csv(combined_predictions, "combined_predictions.csv", row.names = FALSE)
saveRDS(svm_model, "svm_model.rds")
```
